FTLR proof cases

### 2018-03-12 fully-abstract-comp/ftlr-proof-cases.tex
typo and strengthen anti-reduction lemma
changed call semantics, so it does not 'forget' memory not pointed to by the current stack pointer. Also added footnote with explanation in FTLR proof.

\begin{lemma}[$\lro$ closed under target language antireduction]
  For all $\Phi_S$, $\Phi_T$, $\Phi_T'$, $j$, $n$, if
\[
  \Phi_T \steps[j] \Phi_T' \text{ and } \npair{(\Phi_S,\Phi_T')} \in \lro,
\]
then
\[
  \npair{(\Phi_S,\Phi_T)} \in \lro
\]
\end{lemma}

\dominique{9-3-2018: the below is a proposed generalisation of the above antireduction lemma  that I expect will hold (based on experience with similar LRs in the past).}
\begin{lemma}[$\lro$ closed under antireduction (generalised previous lemma)]
  For all $\Phi_S$, $\Phi_S'$, $\Phi_T$, $\Phi_T'$, $j_S,J_T$, $n$, if
  \begin{itemize}
  \item $\Phi_S \steps[j_S] \Phi_S'$
  \item $\Phi_T \steps[j_T] \Phi_T'$
  \item $\npair{(\Phi_S',\Phi_T')} \in \lro$
\end{itemize}
then
\[
  \npair[n+\min(j_S,j_T)]{(\Phi_S,\Phi_T)} \in \lro
\]
\end{lemma}

@@

\section{FTLR proof}
By induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}.
\begin{proof}
  Assume
\[
  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}
  \npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}
\]
and show
\[
\npair{((\rx,\lin),\baddr,\eaddr,\aaddr),
       ((\rx,\lin),\baddr,\eaddr,\aaddr)} \in \lre(W)
\]
We write $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$.\\

@@

\paragraph{Case: $\scall[o]{}{r_1}{r_2}$}
In case $\scall[o]{}{r_1}{r_2}$ is the first step of the execution. We can assume $r_1,r_2 \neq \pcreg$ as this would cause the execution to fail\footnote{{\color{red} TODO:} We still have to consider what happens if the temp register is jumped to.}. By assumption, $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ which in particular means that for suitable $W_{R,1}$ and $W_{R,2}$
\[
  \npair[n']{(\src{\reg_S}(r_i),\reg_T(r_i))} \in \lrv(W_{r,1}) \text{ for i = 1,2}
\]
We can further deduce that $\src{\reg_S}(r_i)$ for $i=1,2$ are sealed capabilities as the execution would otherwise fail. This gives us
\[
  \begin{array}{l}
    \forall W' \future W_{R,1}, W_o, n'' < n', (c_S,c_T) \in H_\sigma \; \sigma \; W_o \ldotp \\
    \quad \npair[n'']{\Phi_S(r_1),c_S,\Phi_T(r_1),\sealed{\sigma,c_T}} \in \lrexj(W' \oplus W_o) \wedge \dots
  \end{array}
\]
and
\[
  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}
\]
At this point, we need to pick a suitable $W' \future W_{R,1}$. To this end, we need to know a bit about $W_{R,1}$. From the operational semantics, we know $\Phi_S(\rstk) = \stkptr{\perm,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\Phi_S(\rstk) \in \lrv(W_{R,\stk})$. From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $W_{R,\stk} \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude. Perhaps a notion of well-formed worlds would be good along with a couple of lemmas that use it to make this reasoning once and for all.}. As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $W_{R,1}(r)$ must be a similar spatial region to the one in $W_{R,\stk}$.
We pick $W'$ to be
\[
  \begin{array}{r l}
    W' = W     & \update{\mathrm{free}.r}{\revoked}\\
               & \update{\mathrm{free}.x}{\stdreg{[\baddr_\stk',\aaddr_\stk -1]}{\spa}}\\
               & \update{\mathrm{priv}.y}{\stdreg{[\aaddr_\stk,\eaddr_\stk']}{\spa}}
  \end{array}
\]
where $x$ and $y$ are fresh region names in the respective part of the world. To indicate similar worlds to $W'$ with the same ownership as another world well-defined w.r.t. $W$, we write $W'_{R,1}$ with the same spatially owned region as $W_{R,1}$. Using the private part of the future world relation, it is easy to see that $W'$ is a future world of $W_{R,1}$ (by the operational semantics $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$). Using world monotonicity of $H_\sigma \; \sigma$\footnote{{\color{red}This means that $H_\sigma$ must be monotone in the world!}}, we get
\[
  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W'_{R,2}
\]
which we use to get:
\[
  \npair[n'']{\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2)} \in \lrexj(W_{R,1}' \oplus W_{R,2}')
\]
In order to use this, we need to pick a register-file pair and suitable memories. We pick a $\Phi_S$ in $\lrexj$, so the $\Phi_S'$ in $\lrexj$ becomes the $\Phi_S'$ we assumed $\Phi_S$ steps to. Specifically
\[
  \Phi_S.\reg
  \begin{array}[t]{l}
    \update{r_1}{\linCons{\Phi_S(r_1)}}\\
    \update{r_2}{\linCons{\Phi_S(r_2)}}\\
    \update{\pcreg}{0} \\
    \update{\rdata}{0} \\
    \update{\rstk}{c_\stk}\\
    \update{\rretc}{c_{\ret,c}}\\
    \update{\rretd}{c_{\ret,d}}\\
    \update{\rtmp{1}}{0}\\
  \end{array}
\]
where
\begin{itemize}
\item $c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1}$
\item $c_{\ret,c} = \retptrc(\sigma',(\perm,\lin),\baddr,\eaddr,\aaddr+\calllen))$
\item $c_{\ret,d} = \retptrd(\sigma',(\aaddr_\stk,\eaddr_\stk))$
\end{itemize}
and
$\ms_S,((\dots,\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}) :: \stk),\ms_\stk|_{[\baddr_\stk,\aaddr_\stk-1]}$

>>>>

\begin{lemma}[$\lro$ closed under target language antireduction]
  For all $\Phi_S$, $\Phi_T$, $\Phi_T'$, $j$, $n$, if
\[
  \Phi_T \nstep[j] \Phi_T' \text{ and } \npair[n-j]{(\Phi_S,\Phi_T')} \in \lro,
\]
then
\[
  \npair{(\Phi_S,\Phi_T)} \in \lro
\]
\end{lemma}

\dominique{9-3-2018: the below is a proposed generalisation of the above antireduction lemma  that I expect will hold (based on experience with similar LRs in the past).}
\begin{lemma}[$\lro$ closed under antireduction (generalised previous lemma)]
  For all $\Phi_S$, $\Phi_S'$, $\Phi_T$, $\Phi_T'$, $j_S,J_T$, $n$, if
  \begin{itemize}
  \item $\Phi_S \nstep[j_S] \Phi_S'$
  \item $\Phi_T \nstep[j_T] \Phi_T'$
  \item $\npair{(\Phi_S',\Phi_T')} \in \lro$
\end{itemize}
then
\[
  \npair[n+\min(j_S,j_T)]{(\Phi_S,\Phi_T)} \in \lro
\]
\end{lemma}

@@

\section{FTLR proof}
By induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}.
\begin{proof}
  Assume
\[
  \begin{array}{l}
  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}\\
  \npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}
  \end{array}
\]
and show
\[
\npair{((\rx,\lin),\baddr,\eaddr,\aaddr),
       ((\rx,\lin),\baddr,\eaddr,\aaddr)} \in \lre(W)
\]
We write $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$.\\

@@

\paragraph{Case: $\scall[o]{}{r_1}{r_2}$}
In case $\scall[o]{}{r_1}{r_2}$ is the first step of the execution. We can assume $r_1,r_2 \neq \pcreg$ as this would cause the execution to fail\footnote{Assuming $\pcreg$ was one of the two registers, then the step relation would reqiure the register to contain an executable capability while the call semantics would require there to be a sealed capability.}\footnote{{\color{red} TODO:} We still have to consider what happens if the temp register is jumped to.}. By assumption, $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ which in particular means that for suitable $W_{R,1}$ and $W_{R,2}$
\[
  \npair[n']{(\src{\reg_S}(r_i),\reg_T(r_i))} \in \lrv(W_{r,1}) \text{ for i = 1,2}
\]
We can further deduce that $\src{\reg_S}(r_i)$ for $i=1,2$ are sealed capabilities as the execution would otherwise fail. This gives us
\[
  \begin{array}{l}
    \forall W' \future W_{R,1}, W_o, n'' < n', (c_S,c_T) \in H_\sigma \; \sigma \; W_o \ldotp \\
    \quad \npair[n'']{\Phi_S(r_1),c_S,\Phi_T(r_1),\sealed{\sigma,c_T}} \in \lrexj(W' \oplus W_o) \wedge \dots
  \end{array}
\]
and
\[
  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}
\]
At this point, we need to pick a suitable $W' \future W_{R,1}$. To this end, we need to know a bit about $W_{R,1}$. From the operational semantics, we know $\Phi_S(\rstk) = \stkptr{\perm,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\Phi_S(\rstk) \in \lrv(W_{R,\stk})$. From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $W_{R,\stk} \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude. Perhaps a notion of well-formed worlds would be good along with a couple of lemmas that use it to make this reasoning once and for all.}. As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $W_{R,1}(r)$ must be a similar spatial region to the one in $W_{R,\stk}$.
We pick $W'$ to be
\[
  \begin{array}{r l}
    W' = W     & \update{\mathrm{free}.r}{\revoked}\\
               & \update{\mathrm{free}.x}{\stdreg{[\baddr_\stk',\aaddr_\stk -1]}{\spa}}\\
               & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk']}{\spa},\opc)}
  \end{array}
\]
\lau{12-05-2018: Write what $\opc$ is.}
where $x$ and $y$ are fresh region names in the respective part of the world. To indicate similar worlds to $W'$ with the same ownership as another world well-defined w.r.t. $W$, we write $W'_{R,1}$ with the same spatially owned region as $W_{R,1}$. Using the private part of the future world relation, it is easy to see that $W'$ is a future world of $W_{R,1}$ (by the operational semantics $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$). Using world monotonicity of $H_\sigma \; \sigma$\footnote{{\color{red}This means that $H_\sigma$ must be monotone in the world!}}, we get
\[
  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W'_{R,2}
\]
which we use to get:
\[
  \npair[n'']{\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2)} \in \lrexj(W_{R,1}' \oplus W_{R,2}')
\]
In order to use this, we need to pick a register-file pair and suitable memories. We pick a $\Phi_S$ in $\lrexj$, so the $\Phi_S'$ in $\lrexj$ becomes the $\Phi_S'$ we assumed $\Phi_S$ steps to. Specifically
\[
  \Phi_S.\reg
  \begin{array}[t]{l}
    \update{r_1}{\linCons{\Phi_S(r_1)}}\\
    \update{r_2}{\linCons{\Phi_S(r_2)}}\\
    \update{\pcreg}{0} \\
    \update{\rdata}{0} \\
    \update{\rstk}{c_\stk}\\
    \update{\rretc}{c_{\ret,c}}\\
    \update{\rretd}{c_{\ret,d}}\\
    \update{\rtmp{1}}{0}\\
  \end{array}
\]
where
\begin{itemize}
\item $c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1}$
\item $c_{\ret,c} = \retptrc(\sigma',(\perm,\lin),\baddr,\eaddr,\aaddr+\calllen))$
\item $c_{\ret,d} = \retptrd(\sigma',(\aaddr_\stk,\eaddr_\stk))$
\end{itemize}
and
$\ms_S,((\opc,\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}) :: \stk),\ms_\stk-\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}$
